---
title: "Lesson 9: Relational data"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<br>

## Readings

#### Required:
[Chapter 13 in 'R for Data Science'](https://r4ds.had.co.nz/relational-data.html) by Hadley Wickham & Garrett Grolemund

Jenny Bryan's [STAT545 Chapter 10: Be the boss of your factors](https://stat545.com/multiple-tibbles.html)

<br>

#### Other resources:
Jenny Bryan's [STAT545 Chapter 14 When one tibble is not enough](https://stat545.com/multiple-tibbles.html)

[Chapter 15 in 'R for Data Science'](https://r4ds.had.co.nz/factors.html) by Hadley Wickham & Garrett Grolemund


## Learning objectives
By the end of today's class, you should be able to:


Re-order the levels of a factor
Re-name and combine the levels of a factor



## Setup
```{r}
library(tidyverse)



```



## Part 1: Relational data
From R4Ds

It’s rare that a data analysis involves only a single table of data. Typically you have many tables of data, and you must combine them to answer the questions that you’re interested in. Collectively, multiple tables of data are called relational data because it is the relations, not just the individual datasets, that are important.

There are four main types of operations that can be done with two tables:

* [**Binding**](https://stat545.com/multiple-tibbles.html#typology-of-data-combination-tasks), which simply stacks tables on top of or beside each other

* [**Mutating joins**](), which add new variables to one data frame from matching observations in another.

* Filtering joins, which filter observations from one data frame based on whether or not they match an observation in the other table.

* Set operations, which treat observations as if they were set elements.

We will only cover the first three today. Let's click on the links to work through the corresponding section in Jenny Bryan's STAT 545 notes or Grolemund and Wickham's R for Data Science.




## Keys


## Types of joins

The most commonly used join is the left join: you use this whenever you look up additional data from another table, because it preserves the original observations even when there isn’t a match. The left join should be your default join: use it unless you have a strong reason to prefer one of the others.

Another way to depict the different types of joins is with a Venn diagram:





## Part 2: Factors in R

In R, factors are used to work with categorical variables, variables that have a fixed and known set of possible values. They are also useful when you want to display character vectors in a non-alphabetical order. The values a factor can take on are called the levels. When working with factors, the two most common operations are changing the order of the levels, and changing the values (names) of the levels. 

To work with factors, we’ll use the `forcats` package, which is part of the core `tidyverse`. It provides tools for dealing with **cat**egorical variables (and it’s an anagram of factors!) using a wide range of helpers for working with factors.

We'll work through [Chapters 15.3-15.6 of R for Data Science](https://r4ds.had.co.nz/factors.html#general-social-survey) to explore some of the functionality.



Exercise from gapminder or flights
