---
title: "Fix the Plots"
output: 
  html_document:
    keep_md: true
    theme: paper
---

<!---The following chunk allows errors when knitting--->

```{r allow errors, echo = FALSE}
knitr::opts_chunk$set(error = TRUE, warning = FALSE)
```

<br>

## Question 1

Why is this not running?

```{r, eval = FALSE}
coronavirus %>%  
  filter(type == "confirmed") %>% 
  group_by(date) %>% 
  summarize(total_cases = sum(cases)) %>% 
  ggplot(mapping = aes(x = date, y = total_cases)) %>%
  geom_line()

```

<br>

## Question 2

What is the problem here?

```{r, eval = FALSE}

top10_countries <- coronavirus %>% 
  filter(type == "confirmed") %>%
  group_by(Country.Region) %>%
  summarize(total_cases = sum(cases)) %>%
  arrange(-total_cases) %>% 
  head(10) %>% 
  .$Country.Region

coronavirus %>%  
  filter(type = confirmed, Country.Region %in% top10_countries) %>% 
  group_by(Country.Region, date) %>% 
  summarize(total_cases = sum(cases)) %>% 
  ggplot(mapping = aes(x = date, y = total_cases, color = Country.Region)) +
  geom_line()

```

<br>

## Question 3

What is the difference between the two code chunks below (look in the `ggplot()` line)? Which would you use and why?

```{r, eval = FALSE}

coronavirus %>%  
  filter(type == "confirmed", Country.Region %in% top10_countries) %>%
  group_by(Country.Region, date) %>% 
  summarize(total_cases = sum(cases)) %>% 
  ggplot(mapping = aes(x = date, y = total_cases, color = Country.Region)) +
  geom_line()

coronavirus %>%  
  filter(type == "confirmed", Country.Region %in% top10_countries) %>%
  group_by(Country.Region, date) %>% 
  summarize(total_cases = sum(cases)) %>% 
  ggplot(aes(date, total_cases, color = Country.Region)) +
  geom_line()

```

<br>

## Question 4

With both the `mutate()` and the `summarize()` functions, we create a new variable using the following syntax: `var_name = function(variable)`. What is the difference between those two functions?


