---
title: "Lesson 6 exercises"
author: "Nina"
date: "4/7/2020"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Exercise

Add a new variable that shows the *proportion of confirmed cases* for which the outcome is still unknown (i.e. not counted as dead or recovered) for each country, and show only countries with more than 20,000 confirmed cases. 

Which country has the lowest proportion of undetermined outcomes? Why?

<br>
<br>

When you're done, sync your RMarkdown file to Github.com (pull, stage, commit, push)

## Exercise

How many countries have suffered more than 3,000 deaths so far, and which three countries have recorded the highest death counts?


## Exercise

Go back to our original dataset `coronavirus` and identify where and when the highest death count in a single day was observed. Hint: you can either use or `base::max` or `dplyr::arrange()`...  
<br>
The first case was confirmed in the US on [January 20 2020](https://www.nejm.org/doi/full/10.1056/NEJMoa2001191), marking the earliest day included in this dataset. When was the first confirmed case recorded in Denmark?

<br>  

<br>  

When you're done, knit your RMarkdown file, and sync it to GitHub (pull, stage, commit, push)


## Exercise
Which day has had the highest total death count globally so far?


## Optional exercise - if you have time
The `month()` function from the package `lubridate` extracts the month from a date. How many countries already have > 1000 deaths in April?


## In-class exercises
1. Which country had the highest number of deaths on Monday (April 6 20202)?
  ```{r, include = FALSE, eval = FALSE}
coronavirus %>% 
  select(-Lat, -Long) %>% 
  filter(date == "2020-04-06", type == "death") %>% 
  arrange(-cases)

```

2. Which country had the highest count of confirmed cases in January? [Hint: to address this question the function `month()` from the package `lubridate` might be helpful]. What about in March?
  ```{r, eval = FALSE, include = FALSE}
library(lubridate) #install.packages('lubridate')

coronavirus %>% 
  mutate(month = month(date)) %>% 
  filter(type == "confirmed", month == 1) %>% 
  group_by(Country.Region) %>% 
  summarize(total_death = sum(cases)) %>% 
  arrange(-total_death)
```


3. Which countries have data for multiple states or provinces?
  ```{r, include = FALSE, eval = FALSE}
coronavirus %>% 
  group_by(Country.Region, date) %>% 
  summarize(n = n()) %>% 
  group_by(Country.Region) %>% 
  summarize(maxcount = max(n)) %>% 
  filter(maxcount > 3)

```

4. Do all countries have reports of the number of confirmed cases for the same number of days?
  ```{r, include = FALSE, eval = FALSE}
coronavirus %>% 
  filter(type == "recovered") %>% 
  group_by(Country.Region, Province.State) %>% 
  summarize(n = n()) %>% 
  arrange(n) %>% 
  head

```

